### 2020-01-15 manuscript/bef_stability/manuscript.Rmd
One PCA + compute clean statistical model

library(tidyverse)
library(magrittr)
library(cowplot)
source(mypath("R", "misc.R"))
source(mypath("R", "plot_methods.R"))
source(mypath("R", "community_analysis.R"))
source(mypath("R", "community_methods.R"))
source(mypath("R", "statistical_analysis.R"))
source(mypath("R", "press_methods.R"))
source(mypath("R", "geo_methods.R"))
source(mypath("R", "synchrony.R"))

@@

```{r}
myload(
  temporal_community_metrics,
  dir = mypath("data")
)

myload( temporal_network_metrics, dir = mypath("data", "classes"))
temporal_network_metrics_classes <- temporal_network_metrics
rm(temporal_network_metrics)

stab_troph <- temporal_network_metrics_classes %>%
  select(station, troph_group) %>%
  unnest(troph_group) %>%
  filter(troph_group != 1) %>%
  mutate(troph_group = as.factor(troph_group)) %>%
  select(station, troph_group, biomass_stab, richness_avg)
```

@@

```{r, fig.show = "hide"}
evt_var_mean <- c("flow_med", "temperature_med", "width_river_mean",
  "avg_depth_station_mean", "slope", "alt", "d_source", "strahler")
evt_var_cv <- c("flow_cv", "temperature_cv", "width_river_cv",
  "avg_depth_station_cv")
evt_var_mean_ok <-  c(evt_var_mean, "DBO_med")
evt_var_cv_ok <-  c(evt_var_cv, "DBO_cv")
pca_pressure <- habitat_pressure[, names(habitat_pressure) %in% c(evt_var_mean_ok, evt_var_cv_ok, "station")] %>%
  na.omit

pca_list <- map(list(evt_var_mean_ok, evt_var_cv_ok), function (.var) {
  pca_pressure[, names(pca_pressure) %in% .var]
})
names(pca_list) <- c("evt_med", "evt_cv")

pca_rotated <- compute_rotated_pca(
  data_list = pca_list,
  titles = names(pca_list)  
)

pca_rotated_plot <- plot_rotated_pca(pca_rotated = pca_rotated)
names(pca_rotated_plot)
```

@@

```{r}
stab_div_mod <- lme(data = sem_data,
  fixed = biomass_stab ~ richness_med,
  random = ~ 1 | basin/(alt + width)
)
rsquared(stab_div_mod, method = "nagelkerke")

@@

library(ggraph)
p_stab_sem <- ggpiecewisesem(stab_sem)
p_bm_sem <- ggpiecewisesem(stab_sem)
```

@@

```{r s-div, fig.dim = c(7,3)}
# Plot left  
p_stab_div <- temporal_community_metrics %>%
  ggplot(aes(x = richness_med, y = biomass_stab))+
  geom_smooth(method = "lm") +
  geom_point() +
  labs(x = "Median species richness", y = "Biomass stability")

# Right plot
p_stab_div_troph <- stab_div_troph %>%
  ggplot(aes(x = richness_avg, y = biomass_stab)) +
  geom_point(aes(color = troph_group)) +
  geom_smooth(method = "lm") +
  ggplot2::scale_color_manual(
    labels = troph_group_labeller(),
    values = c("blue", "green", "red"),
    name = "Trophic group"
  ) +
  labs(x = "Average species richness", y = "Biomass stability")

@@

```{r sem, fig.dim = c(10,6)}
#https://tex.stackexchange.com/questions/20496/drawing-different-tikz-shapes-parameterized-by-data-from-a-file?rq=1
#https://tex.stackexchange.com/questions/83888/how-to-plot-data-from-a-csv-file-using-tikz-and-csvsimple
plot_grid(
  plot_grid(pca_rotated_plot$evt_med$rotated,
  pca_rotated_plot$evt_cv$rotated),
  p_stab_sem,
  p_bm_sem, labels = "AUTO"
)
```

- Habitat:
  - Median:
    - Axis1 (river width, dist source, strahler)
    - Axis2 (DBO, slope)
  - CV: 
    - Axis1 (temperature cv, flow cv)
    - Axis2 (width)

>>>>

library(tidyverse)
library(magrittr)
library(ggeffects)
library(cowplot)
source(mypath("R", "misc.R"))
source(mypath("R", "plot_methods.R"))
source(mypath("R", "community_analysis.R"))
source(mypath("R", "community_methods.R"))
source(mypath("R", "statistical_analysis.R"))
source(mypath("R", "press_methods.R"))
source(mypath("R", "geo_methods.R"))
source(mypath("R", "synchrony.R"))

@@

```{r}
myload(temporal_community_metrics, dir = mypath("data"))

myload(temporal_network_metrics, dir = mypath("data", "classes"))
temporal_network_metrics_classes <- temporal_network_metrics
myload(temporal_network_metrics, dir = mypath("data", "species"))

stab_div_troph <- temporal_network_metrics %>%
  dplyr::select(station, troph_group) %>%
  unnest(troph_group) %>%
  filter(troph_group != 1) %>%
  mutate(troph_group = as.factor(troph_group)) %>%
  dplyr::select(station, troph_group, biomass_stab, richness_med)
```

@@

```{r, fig.show = "hide"}
evt_var_mean <- c("flow_med", "temperature_med", "width_river_mean",
  "avg_depth_station_mean", "slope", "alt", "d_source", "strahler", "DBO_med")
evt_var_cv <- c("flow_cv", "temperature_cv", "width_river_cv",
  "avg_depth_station_cv", "DBO_cv")
pca_pressure <- habitat_pressure[, names(habitat_pressure) %in% c(evt_var_mean, evt_var_cv, "station")] %>%
  na.omit

pca_rotated <- compute_rotated_pca(
  .data = pca_pressure[, !names(pca_pressure) %in% "station"],
  naxis = 5 
)

@@

$$
S_i = \beta_0 + \beta_Rx_{Ri} + \epsilon_{ij}
$$

```{r mod-div-total}
ctrl <- lmeControl(opt='optim')
stab_div_mod <- lme(data = sem_data,
  fixed = log_stab ~ log_rich,
  random = ~ 1 + log_rich | basin,
  control = ctrl 
)
#stab_div_colin <- car::vif(stab_div_mod)
stab_div_sum <- piecewiseSEM::rsquared(stab_div_mod, method = "nagelkerke")
```

```{r predict}
pred <- ggpredict(stab_div_mod, terms = c("log_rich"), type = "fe")
stab_div_pred <- pred %>%
  as_tibble() %>%
  rename(
    log_stab = predicted,
    log_rich = x,
    basin = group
  )

@@

library(ggraph)
p_stab_sem <- ggpiecewisesem(stab_sem)
p_bm_sem <- ggpiecewisesem(bm_sem)
```

@@

```{r s-div, fig.dim = c(10,5)}
label_rsq <- "paste(
  \"Conditional italic(R) ^ 2 == \", round(stab_div_sum$Conditional, 2),
  \", Marginal italic(R) ^ 2 == \", round(stab_div_sum$Marginal, 2))"
# Plot left  
p_stab_div <- sem_data %>%
  ggplot(aes(x = log_rich, y = log_stab))+
  geom_point() +
  #scale_x_log10() + scale_y_log10() +
  labs(x = "Median species richness (Log 10)", y = " Biomass stability (Log 10)") +
  geom_line(data = stab_div_pred) +
  geom_ribbon(data = stab_div_pred, aes(ymin = conf.low, ymax = conf.high), fill
  = "grey20", alpha = .5) +
  annotate("text", x = 1.0, y = .8, label = paste(
  " Marg. R^2 = ", round(stab_div_sum$Marginal, 2),
  ", Cond. R^2 = ", round(stab_div_sum$Conditional, 2)
  )
    )

# Right plot
p_stab_div_troph <- stab_div_troph %>%
  ggplot(aes(x = richness_med, y = biomass_stab, color = troph_group)) +
  geom_point() +
  geom_smooth(method = "lm" ) +
  ggplot2::scale_color_manual(
    labels = troph_group_labeller(),
    values = c("blue", "green", "red"),
    name = "Trophic group"
  ) +
  scale_x_log10() + scale_y_log10() +
  labs(x = "Average species richness", y = "Biomass stability")

@@

```{r, fig.show = "hide"}
axis_pair <- list(c(1,2), c(1,3), c(2,3), c(1, 4), c(2, 4), c(3,4), c(4,5))
p_l_rot <- map(axis_pair, 
  function (axis)
    pca_rotated_plot <- plot_rotated_pca(
      pca_rotated = pca_rotated,
      axis = axis 
    )
)
```

```{r sem-plot, fig.dim = c(10,10)}
#https://tex.stackexchange.com/questions/20496/drawing-different-tikz-shapes-parameterized-by-data-from-a-file?rq=1
#https://tex.stackexchange.com/questions/83888/how-to-plot-data-from-a-csv-file-using-tikz-and-csvsimple
to_plot <- c(1, 6, 7)
plot_grid(
  plot_grid(plotlist = map(p_l_rot[to_plot], ~.x$rotated),
      ncol = length(to_plot), labels = LETTERS[1]),
  plot_grid(p_stab_sem, p_bm_sem, labels = LETTERS[2:3]),
  ncol = 1,
  rel_heights = c(.4, .6)
)
```

- Habitat: (+ / -)
  - RC 1: width & depth & flow med, d source, strahler 	/
  - RC 2: T°C med 					/ slope, alt 
  - RC 3: 						/ flow CV, DBO med
  - RC 4: river width med, depth med 			/
  - RC 5: DBO CV 					/

### 2020-01-15 report/02-methods.Rmd
One PCA + compute clean statistical model

#### Separate Med and CV pressure

```{r pca-med-cv, fig.cap="PCA with Avg or CV variables"}
hab_med <- habitat_pressure %>%
  select_at(vars(matches("med|mean")))
hab_cv <- habitat_pressure %>%
  select_at(vars(matches("_cv"))) %>%
  select(-width_river_cv, -avg_depth_station_cv, -nitrates_cv, -phosphore_cv)

plot_pca <- function (.data, title = NULL) {

  if (is.null(title)){
  title <- paste0("PCA - ",
      ifelse(any(!is.na(str_match(names(.data), "_cv"))), "CV",
      "Median"))
  }

  pca_hab <- dudi.pca(as.data.frame(na.omit(.data)), scannf = FALSE, nf = 3, center = TRUE, scale = TRUE) 
  eig_plot <- fviz_eig(pca_hab)
  pca_plot <- fviz_pca_var(pca_hab,
    axes = c(1, 2),
    col.var = "contrib", # Color by contributions to the PC
    gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
    title = title,
    repel = TRUE # Avoid text overlapping
  )
  pca_plot12 <- ggdraw(pca_plot) +
    draw_plot(eig_plot, x = .7, y = .95, hjust = 1, vjust = 1, width = 0.17, height = 0.25)
  pca_plot12
  }
plot_pca_med_cv <- map(list(hab_med, hab_cv), plot_pca)
plot_grid(plotlist = plot_pca_med_cv, labels = "AUTO")
```

#### Separate according to physical habitat, climate, enrichment

```{r pca-evt}
phys_var <- c( "width_river_mean", "avg_depth_station_mean",
   "slope", "alt", "d_source", "strahler")
#"width_river_cv", "avg_depth_station_cv", "under_bank_med", "hole_pit_med",
#"long", "lat","logjam_stumps_med", "edge_vg_med",
#"rock_shelter_med", "aqua_vg_shelter_med", "shade_med", "sinuosite_med",
climate_var <- c("flow_med", "temperature_med", "flow_cv", "temperature_cv") 
pressure_var <- c( "fungicides_cv", "herbicides_cv",
  "insecticides_cv",  "fungicides_med",
  "herbicides_med", "insecticides_med")
#, "nitrates_med", "phosphore_med","nitrates_cv", "phosphore_cv","DBO_med", "DBO_cv"
hab_physical <- habitat_pressure[, names(habitat_pressure) %in% phys_var]
hab_climate <- habitat_pressure[, names(habitat_pressure) %in% climate_var]
hab_pressure <- habitat_pressure[, names(habitat_pressure) %in% pressure_var]

p_pca_clim_phy_press <- map2(
  list(hab_physical, hab_climate, hab_pressure),
  c("Physical", "Climate", "Pressure"),
  plot_pca)
plot_grid(plotlist = p_pca_clim_phy_press, labels = "AUTO")
```

####

```{r pca-evt-press-med-cv}
evt_var_mean <- c("flow_med", "temperature_med", "width_river_mean",
  "avg_depth_station_mean", "slope", "alt", "d_source", "strahler")
evt_var_cv <- c("flow_cv", "temperature_cv", "width_river_cv",
  "avg_depth_station_cv")

press_var_cv <- c("fungicides_cv", "herbicides_cv", "insecticides_cv", "DBO_cv")
press_var_med <- c("fungicides_med", "herbicides_med",
  "insecticides_med","DBO_med")

#, "nitrates_med", "phosphore_med","nitrates_cv", "phosphore_cv", "DBO_cv"
hab_evt <- habitat_pressure[, names(habitat_pressure) %in% c(evt_var_mean,
  evt_var_cv)]
hab_press <- habitat_pressure[, names(habitat_pressure) %in% c(press_var_med,
  press_var_cv)]

p_pca_clim_phy_press <- map2(
  list(hab_evt, hab_press),
  c("EVT", "Pressure"),
  plot_pca)
plot_grid(plotlist = p_pca_clim_phy_press, labels = "AUTO")
```

####

```{r}
hab_mean <- habitat_pressure[, names(habitat_pressure) %in% c(evt_var_mean , press_var_med)]
hab_cv <- habitat_pressure[, names(habitat_pressure) %in% c(evt_var_cv, press_var_cv)]

p_pca_clim_phy_press <- map2(
  list(hab_mean, hab_cv),
  c("Med", "Cv"),
  plot_pca)
plot_grid(plotlist = p_pca_clim_phy_press, labels = "AUTO")
```

####

```{r}
pca_pressure <- habitat_pressure[, names(habitat_pressure) %in% c(evt_var_mean, evt_var_cv, press_var_med, press_var_cv, "station")] %>% 
  na.omit
hab_evt_mean <- pca_pressure[, names(pca_pressure) %in% evt_var_mean]
hab_evt_cv <- pca_pressure[, names(pca_pressure) %in% evt_var_cv]
hab_press_mean <- pca_pressure[, names(pca_pressure) %in% press_var_med]
hab_press_cv <- pca_pressure[, names(pca_pressure) %in% press_var_cv]

p_pca_clim_phy_press <- map2(
  list(hab_evt_mean, hab_evt_cv, hab_press_mean, hab_press_cv),
  c("Med evt", "Cv evt", "Med press", "Cv press"),
  plot_pca)
plot_grid(plotlist = p_pca_clim_phy_press, labels = "AUTO")
```

```{r}
pca_clim_phy_press <- map(
  list(hab_evt_mean, hab_evt_cv, hab_press_mean, hab_press_cv),
  function(.data) dudi.pca(as.data.frame(na.omit(.data)), scannf = FALSE, nf = 3, center = TRUE, scale = TRUE))
names(pca_clim_phy_press) <- c("med_evt", "cv_evt", "med_press", "cv_press")
```

#### Rotate PCA axis

```{r, fig.show = "hide"}
pca_rotated <- compute_rotated_pca(
  data_list = list(hab_evt_mean, hab_evt_cv, hab_press_mean, hab_press_cv),
  titles = c("med_evt", "cv_evt", "med_press", "cv_press")
)

test <- plot_rotated_pca(pca_rotated)

mysave(pca_rotated, dir = mypath("report"), overwrite = TRUE)
```

```{r pca_rotated_plot}
plot_grid(plotlist = test)
```

#### Rotate PCA axis 2

```{r pca-data, fig.show = "hide"}
evt_var_mean_ok <-  c(evt_var_mean, "DBO_med")
evt_var_cv_ok <-  c(evt_var_cv, "DBO_cv")
pca_pressure <- habitat_pressure[, names(habitat_pressure) %in% c(evt_var_mean_ok, evt_var_cv_ok, "station")] %>% 
  na.omit

pca_list <- map(list(evt_var_mean_ok, evt_var_cv_ok), function (.var) {
  pca_pressure[, names(pca_pressure) %in% .var]
})
names(pca_list) <- c("evt_med", "evt_cv")

pca_rotated <- compute_rotated_pca(
  data_list = pca_list,
  titles = names(pca_list)  
)

pca_rotated_plot <- plot_rotated_pca(pca_rotated = pca_rotated)
```

```{r pca_env_ok}
plot_grid(plotlist = pca_rotated_plot, ncol = 1)
```

@@

test <- map(pca_rotated, function(.pca) .pca$rotated$scores)
axis_names <- paste0(rep(names(pca_rotated), each = 2), c("1", "2"))
test2 <- do.call(cbind, test)
colnames(test2) <- axis_names
rotated_axis_st <- cbind(pca_pressure[, "station"], test2)

>>>>

#### 1 rotated PCA

```{r}
evt_var_mean <- c("flow_med", "temperature_med", "width_river_mean",
  "avg_depth_station_mean", "slope", "alt", "d_source", "strahler", "DBO_med")
evt_var_cv <- c("flow_cv", "temperature_cv", "width_river_cv",
  "avg_depth_station_cv", "DBO_cv")
pca_pressure <- habitat_pressure[, names(habitat_pressure) %in% c(evt_var_mean, evt_var_cv, "station")] %>%
  na.omit

pca_rotated <- compute_rotated_pca(
  .data = pca_pressure[, !names(pca_pressure) %in% "station"],
  naxis = 5 
)

axis_pair <- list(c(1,2), c(1,3), c(2,3), c(1, 4), c(2, 4), c(3,4), c(4,5))
p_l_rot <- map(axis_pair, 
  function (axis)
    pca_rotated_plot <- plot_rotated_pca(
      pca_rotated = pca_rotated,
      axis = axis 
    )
)
to_plot <- c(1, 6, 7)
plot_grid(plotlist = c(
    map(p_l_rot[to_plot], ~.x$rotated),
    map(p_l_rot[to_plot], ~.x$normal)
    ), ncol = length(to_plot))
```

@@

test <- pca_rotated$rotated$scores
rotated_axis_st <- cbind(pca_pressure[, "station"], test)
